---
title: ""
author: "Antonio Checa"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document :
      css: IMP_ReportCSS.css
      code_folding: hide
      number_sections: false
      toc: true
      toc_float: 
        collapsed: false
        smooth_scroll: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(knitr)
library(tidyverse)
library(DT)
library(htmltools)
library(gridExtra)
library(writexl)


# set default options
opts_chunk$set(echo = FALSE,
               message = FALSE,
               fig.align="center",
               fig.width = 5,
               fig.height = 5,
               dpi = 120,
               fig.retina = 3)

```

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri(file.path("imp_logo.jpg")), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:70px;')
```




# 1. Aim (s)

Inspection of data quality for acquired SL data in plasma for project SMMSCF-34
Inspecting of peaks passing QC
Creation of final file to report results


# 2. Data import and summary

- Data was imported into R. 

- The final number of samples per group was summarized.


```{r import, echo = TRUE}

importData <- readRDS(file = "Data/File_SamplesSelected_Set.RDS")

samples_file_notreplaced <- importData$samples_file_notreplaced 
samples_file_replaced <- importData$samples_file_replaced 

compound_names <- importData$compound_names #Keep names from previous analysis

compoundInfo <- importData$compoundInfo
rm(importData)

# Summarize injections (Excluded calibration curves and conditioning)



DT::datatable(samples_file_replaced %>% 
  group_by(Sample_type) %>%
  summarize("Number of samples" = n()))

DT::datatable(samples_file_replaced %>% 
  group_by(Sample_type, cohort) %>%
  summarize("Number of samples" = n()))


```




# 4. Quality Summary {.tabset}

A graph is performed per compound

- Individual plots of quantified sphingolipids in plaque

- Allows to inspect instrumental variations in the quantification of the individual species.

- The QC of pooled plasma is a pooled mix of all samples

- %CV for QCs are reported on a table

- Graphs are scaled to include 1.3 x 99% of the samples for visualization purposes

- Values represent concentrations of the compounds already normalized by the extraction volume

- QC does not represent the samples, so values are not representative of the real variance


## 4.1 Quality summary

 
```{r qcnonfilter, echo = TRUE, fig.width= 10, fig.height= 6}

#Summary of CVExt Serum

DT::datatable(samples_file_replaced %>% 
                filter(Sample_type == "QCExt", SMMS_ID != 46) %>%
                pivot_longer(compound_names[,2], names_to = "Compound", values_to = "Conc") %>%
                group_by(Compound) %>%  
                summarize("%CV" = round(sd(Conc, na.rm = TRUE)/mean(Conc, na.rm = TRUE)*100,1)),
                caption = "Plasma. Summary of %CV QCExt by Compound, All Included")



```


## 4.2 SL plots Plaque not normalized Samples and QC

```{r plots GOMall, results="asis", echo = TRUE, warning=FALSE, fig.width=10, fig.height=5}

#Plot compounds by injection order

#Generate plots


  injordplot <- samples_file_replaced %>% filter(Sample_type == "Sample" | Sample_type == "QCExt" 
                                         | Sample_type == "BLExt") %>%
                                             filter(Injection_order != 463) %>% 
                                          arrange(desc(Sample_type))
                                    
  

for (i in compound_names[,2])   {
  
  

  compound <- i
  
  
  axis.graphtitle.text <- element_text(face = "bold", color = "black", size = 12) #Label style
  axis.graph.text <- element_text(face = "bold", color = "black", size = 8)
  
  axis.limit.value <- quantile(injordplot[,i], prob = 0.95, na.rm=TRUE)*1.2 # 40% extra limit for each compound. Quantiles are used to have most of the compounds on the graph. Extreme outliers may be out.
  
    axis.limit.valueq75 <- quantile(injordplot[,i], prob = 0.95, na.rm=TRUE)*1.2 # 20% extra limit for each compound. Quantiles are used to focus on the most common detected values.

  
  #Create graph Sample ID
  
  
p <- ggplot(injordplot, aes(`Injection_order`, unlist(injordplot[,i]))) +
              theme_classic() +   
              geom_point(aes(color = Sample_type), size = 3, alpha = 0.6) +
              scale_colour_manual(values = c("Black","Blue","Orange")) +
           #   ylim(c(0,axis.limit.value)) +
              labs(x="Injection order", y = compound) + 
              theme(legend.position="bottom", axis.text.x=element_text(angle = 0, hjust = 0.5, vjust = 0), 
                    axis.title = axis.graphtitle.text, axis.text = axis.graph.text)


print(i)  

plot(p)

}




```


----------


## 5.1 Sphingolipid plots

- All reported compounds plots per cohort

## 5.1 Samples by Diagnosis

- Plots per Diagnosis Not normalized and Normalized

```{r plots GOMDia, results="asis", echo = TRUE, warning=FALSE, fig.width=6, fig.height=8}


#Plot compounds by injection order

#Generate plots


for (i in compound_names[,2])   {
  
  
  injordplot <- samples_file_replaced %>% filter(Sample_type == "Sample", Replicated == "R01")
  
  compound <- i
  
  axis.graphtitle.text <- element_text(face = "bold", color = "black", size = 12) #Label style
  axis.graph.text <- element_text(face = "bold", color = "black", size = 8)
  
  axis.limit.value <- quantile(injordplot[,i], prob = 1, na.rm=TRUE)*1.3 # 40% extra limit for each compound. Quantiles are used to have most of the compounds on the graph. Extreme outliers may be out.

  
  #Create graph Sample ID
  
  
p <- ggplot(injordplot, aes(`cohort`, unlist(injordplot[,i]))) +
              theme_classic() +   
              scale_colour_manual(values = c("Grey", "Red", "Blue", "Orange")) +
              scale_fill_manual(values = c("Grey", "Red", "Blue", "Orange")) +
              geom_boxplot(aes(fill = cohort), alpha = 0.1) +
              geom_point(aes(color = cohort), size = 4, alpha = 0.5) +
              ylim(c(0,axis.limit.value)) +
              labs(x="Cohort", y = compound) + 
              theme(legend.position="bottom", axis.text.x=element_text(angle = 0, hjust = 0.5, vjust = 0), 
                    axis.title = axis.graphtitle.text, axis.text = axis.graph.text) 

  




print(i)  

grid.arrange(p, nrow=1)
  
}


```



# 6. Selection of samples

Finally, samples are selected as a combination of confirmed detection, presence in blanks and presence in at least 5 samples.

Different Tables are Exported to be included in the final report.



```{r}


# Select compounds to be included. Includes final filter by %CV and technical judgement

compoundInfo_final<- compoundInfo %>% filter(Report == "YES") %>%
                              select(-"Integration",-"our.coments",-"Included", -"Detected",
                                      -"Include",-"Report",-"Use_type")


compoundInfo_excluded <- compoundInfo %>% filter(Included == "YES", Report == "NO") %>% 
                                       select(-"Integration",-"our.coments",-"Included", -"Detected",
                                      -"Include",-"Report",-"Use_type")

compound_excluded <- compoundInfo_excluded %>% select(Compound_name) %>% as.vector()

compound_included<- compoundInfo_final %>% select(Compound_name) %>% as.vector()

smetadata <- colnames(samples_file_replaced [1:35])

finalfile_names_plasma <- samples_file_replaced %>% select(smetadata, compound_included[,1])



# Export files to merge in a the final report.
# This can be done with xlsx package for direct organization in Sheets but there are troubles with Java and Biomedicum computers
# Doing it with write_xlsx will be more practical (the files are edited later anyway)

write_xlsx(finalfile_names_plasma, "FinalResults/ResultsFile_Replaced.xlsx") # The final datatable

write_xlsx(compoundInfo_final, "FinalResults/ReportedCompoundMetadata_Replaced.xlsx") 
write_xlsx(compoundInfo_excluded, "FinalResults/ReportedCompoundMetadataEx_Replaced.xlsx") 

#Save file for future R data processing

list_selected_data <- list(finalfile_names_plasma = finalfile_names_plasma,
                           compoundInfo_final = compoundInfo_final,
                           compoundInfo_excluded = compoundInfo_excluded)

saveRDS(list_selected_data, file = "FinalResults/FinalData_Set_SL_Replaced.RDS")


```


# 7. Final considerations {.tabset}

List of compounds reported and excluded from the final results

## 7.1. Compounds included

```{r}

DT::datatable(compoundInfo_final,
  caption = "Summary of Compounds included")

```

## 7.2 Compounds excluded

```{r}

DT::datatable(compoundInfo_excluded,
  caption = "Summary of Compounds included")

```



# 8. Session Information

Packages used for the creation of this file

```{r session, echo = TRUE, warning=FALSE}

sessionInfo()

```


