---
title: "Quality control report"
author: "Antonio Checa and Matt Smith"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
    html_document :
      css: IMP_ReportCSS.css
      code_folding: hide
      number_sections: false
      toc: true
      toc_float: 
        collapsed: false
        smooth_scroll: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# set default options
knitr::opts_chunk$set(echo = FALSE,
               message = FALSE,
               warning = FALSE,
               fig.align="center",
               fig.width = 5,
               fig.height = 5,
               dpi = 120,
               fig.retina = 3)

```

```{r, echo=FALSE}
htmltools::img(src = knitr::image_uri(sprintf("%s/vignettes/imp_logo.jpg", base_dir)), 
               alt = 'logo', 
               style = 'position:absolute; top:0; right:0; padding:70px;')
```


# 1. Aim (s)

Inspection of data quality for acquired targeted lipidomics data.
Inspecting of peaks passing QC
Creation of final file to report results


# 2. Data import and summary

- Data was imported into R. 

- The final number of samples per group was summarized.


```{r import, echo = FALSE}

importData <- readRDS(file = out_RDS)

metadata <- importData$sample_meta
data_matrix <- importData$data

compoundInfo <- importData$variable_meta
compound_names <- compoundInfo$Compound

# Summarize injections (Excluded calibration curves and conditioning)

DT::datatable(metadata %>% 
  group_by_(.dots=`inc_factor`) %>%
  summarize("Number of samples" = n()))

DT::datatable(metadata %>% 
  group_by_(.dots=c(`inc_factor`, `factor_bps1`, `factor_bps2`)) %>%
  summarize("Number of samples" = n()))


```

# 3. Quality Summary {.tabset}

A graph is performed per compound

- Individual plots of quantified sphingolipids in plaque

- Allows to inspect instrumental variations in the quantification of the individual species.

- The QC of pooled plasma is a pooled mix of all samples

- %CV for QCs are reported on a table

- Graphs are scaled to include 1.3 x 99% of the samples for visualization purposes

- Values represent concentrations of the compounds already normalized by the extraction volume

- QC does not represent the samples, so values are not representative of the real variance


## 3.1 Quality summary

```{r qcnonfilter, echo = FALSE, fig.width= 10, fig.height= 6}

#Summary
filter_wf = 
    filter_smeta(mode = "include",
               levels = qc_name, 
               factor_name = qc_factor)

QC_data =  predicted(model_apply(filter_wf, importData))

QC_data = QC_data$data %>% 
  mutate(name = QC_data$sample_meta[[qc_id_header]]) %>%
  tidyr::pivot_longer(QC_data$variable_meta$Compound, names_to = "Compound", values_to = "value") %>%
  group_by(name, Compound) %>%  
  summarize("%CV" = round(sd(value, na.rm = TRUE)/mean(value, na.rm = TRUE)*100,1))


DT::datatable(QC_data,
                caption = "Summary of %CV by Compound, All Included")

```


# 4. Exploratory analysis {.tabset}

PCA analysis of data to check structure and guide which features are of interest.

Pre-processing

- 50% feature filter

- 50% sample filter

- knn missing value imputation

- pareto scale

- mean centre

## 4.0

Scroll Right!!!

## 4.1 PCA showing `r pca_factor2`

- General PCA

```{r, results="asis", echo = FALSE, include=T, warning=FALSE, fig.width=6, fig.height=5}

# https://bioconductor.org/packages/release/bioc/vignettes/structToolbox/inst/doc/data_analysis_omics_using_the_structtoolbox.html

# Set up peak matrix processing method and PCA.
pmp_wf =
  filter_smeta(mode = "include",
               levels = inc_levels_qc,
               factor_name = inc_factor_qc) +
  mv_feature_filter(threshold = 49,
                    method = "QC",
                    factor_name = feature_filter_factor,
                    qc_label = feature_filter_levels) +
  mv_sample_filter(mv_threshold = 49) +
  mv_feature_filter(threshold = 49,
                    method = "across",
                    factor_name = inc_factor_qc) +
  filter_smeta(mode = "include",
               levels = inc_levels,
               factor_name = inc_factor) +
  structToolbox::knn_impute(neighbours=3,
                            sample_max=50,
                            feature_max=50) +
  structToolbox::pareto_scale() +
  #structToolbox::log_transform() +
  structToolbox::mean_centre()+
  structToolbox::PCA(number_components=4)

# Apply the peak matrix processing and PCA method
general_pmp_wf = model_apply(pmp_wf, combined_datamatrices)

# Plot PCA score 1
scores_p = pca_scores_plot(factor_name=c(pca_factor1),
                           points_to_label = 'none')
chart_plot(scores_p,general_pmp_wf[length(general_pmp_wf)])

# Plot PCA score 2
scores_p = pca_scores_plot(factor_name=c(pca_factor2),
                           points_to_label = 'none')
chart_plot(scores_p,general_pmp_wf[length(general_pmp_wf)])

# Plot PCA loadings
loadings_df = general_pmp_wf[length(general_pmp_wf)]@loadings %>%
  tibble::rownames_to_column("feature")

ggplot(loadings_df, aes(x=PC1, y=PC2)) +
  geom_point(size=2, alpha=1, shape=23) +
  theme_Publication() +
  stat_ellipse(type='norm', level=0.9) +
  ggrepel::geom_text_repel(size=2.5, label.padding = unit(0.2, "lines"),
                           col="red", aes(label=feature), nudge_x = 0, nudge_y = 0.01)
```

## 4.2 PCA subset by `r pca_factor1` {.tabset .tabset-fade}

- PCA subset by levels of the argument 'pca_factor1'

```{r, results="asis", echo = FALSE, include=T, warning=FALSE, fig.width=6, fig.height=5}
# https://bioconductor.org/packages/release/bioc/vignettes/structToolbox/inst/doc/data_analysis_omics_using_the_structtoolbox.html

levs = unique(predicted(general_pmp_wf[length(general_pmp_wf)])$sample_meta[[pca_factor1]])

for(lev in levs){
  pmp_wf =
    filter_smeta(mode = "include",
                 levels = inc_levels_qc,
                 factor_name = inc_factor_qc) +
    filter_smeta(mode = "exclude",
                 levels = as.character(levs[which(levs != lev)]),
                 factor_name = pca_factor1) +
    mv_feature_filter(threshold = 49,
                      method = "QC",
                      factor_name = feature_filter_factor,
                      qc_label = feature_filter_levels) +
    mv_sample_filter(mv_threshold = 49) +
    mv_feature_filter(threshold = 49,
                      method = "across",
                      factor_name = inc_factor_qc) +
    filter_smeta(mode = "include",
                 levels = inc_levels,
                 factor_name = inc_factor) +
    structToolbox::knn_impute(neighbours=3,
                              sample_max=50,
                              feature_max=50) +
    structToolbox::pareto_scale() +
    #structToolbox::log_transform() +
    structToolbox::mean_centre()+
    structToolbox::PCA(number_components=4)
  
  # Apply the peak matrix processing and PCA method
  apply_pmp_wf = model_apply(pmp_wf, combined_datamatrices)
  
  # Plot PCA score 2
  scores_p = pca_scores_plot(factor_name=c(pca_factor2),
                             points_to_label = 'none')
  p = chart_plot(scores_p,apply_pmp_wf[length(apply_pmp_wf)]) + theme(legend.position = "right", legend.direction = "vertical")
  
  # Plot PCA loadings
  loadings_df = apply_pmp_wf[length(apply_pmp_wf)]@loadings %>%
    tibble::rownames_to_column("feature")
  
  p1 = ggplot(loadings_df, aes(x=PC1, y=PC2)) +
    geom_point(size=2, alpha=1, shape=23) +
    theme_Publication() +
    stat_ellipse(type='norm', level=0.9) +
    ggrepel::geom_text_repel(size=2.5, label.padding = unit(0.2, "lines"),
                             col="red", aes(label=feature), nudge_x = 0, nudge_y = 0.01)
  
  leg <- cowplot::get_legend(p)
  
  ###### PLOT
  cat( paste('### ', lev, '\n' ) )
  
  grid.arrange(p + theme(legend.position = "none"),
               p1 + theme(legend.position = "none"),
               leg,
               layout_matrix = rbind(c(1, 1, 1, 3),
                                     c(2, 2, 2, 3)),
               top = sprintf("%s", lev))
  
  cat('\n \n')
  
}

```

## 4.3 PCA subset by `r pca_factor2` {.tabset .tabset-fade}

- PCA subset by levels of the argument 'pca_factor2'

```{r, results="asis", echo = FALSE, include=T, warning=FALSE, fig.width=6, fig.height=5}
# https://bioconductor.org/packages/release/bioc/vignettes/structToolbox/inst/doc/data_analysis_omics_using_the_structtoolbox.html

levs = unique(predicted(general_pmp_wf[length(general_pmp_wf)])$sample_meta[[pca_factor2]])

for(lev in levs){
  pmp_wf =
    filter_smeta(mode = "include",
                 levels = inc_levels_qc,
                 factor_name = inc_factor_qc) +
    filter_smeta(mode = "exclude",
                 levels = as.character(levs[which(levs != lev)]),
                 factor_name = pca_factor2) +
    mv_feature_filter(threshold = 49,
                      method = "QC",
                      factor_name = feature_filter_factor,
                      qc_label = feature_filter_levels) +
    mv_sample_filter(mv_threshold = 49) +
    mv_feature_filter(threshold = 49,
                      method = "across",
                      factor_name = inc_factor_qc) +
    filter_smeta(mode = "include",
                 levels = inc_levels,
                 factor_name = inc_factor) +
    structToolbox::knn_impute(neighbours=3,
                              sample_max=50,
                              feature_max=50) +
    structToolbox::pareto_scale() +
    #structToolbox::log_transform() +
    structToolbox::mean_centre()+
    structToolbox::PCA(number_components=4)
  
  # Apply the peak matrix processing and PCA method
  apply_pmp_wf = model_apply(pmp_wf, combined_datamatrices)
  
  # Plot PCA score 2
  scores_p = pca_scores_plot(factor_name=c(pca_factor1),
                             points_to_label = 'none')
  p = chart_plot(scores_p,apply_pmp_wf[length(apply_pmp_wf)]) + theme(legend.position = "right", legend.direction = "vertical")
  
  # Plot PCA loadings
  loadings_df = apply_pmp_wf[length(apply_pmp_wf)]@loadings %>%
    tibble::rownames_to_column("feature")
  
  p1 = ggplot(loadings_df, aes(x=PC1, y=PC2)) +
    geom_point(size=2, alpha=1, shape=23) +
    theme_Publication() +
    stat_ellipse(type='norm', level=0.9) +
    ggrepel::geom_text_repel(size=2.5, label.padding = unit(0.2, "lines"),
                             col="red", aes(label=feature), nudge_x = 0, nudge_y = 0.01)
  
  leg <- cowplot::get_legend(p)
  
  
  
  ###### PLOT
  cat( paste('### ', lev, '\n' ) )
  
  grid.arrange(p + theme(legend.position = "none"),
             p1 + theme(legend.position = "none"),
             leg,
             layout_matrix = rbind(c(1, 1, 1, 3),
                                   c(2, 2, 2, 3)),
             top = sprintf("%s", lev))
  
  cat('\n \n')
  
}

```


# 5. Individual lipid analysis {.tabset}

All reported compounds plotted based on a factor

## 5.0

Scroll Right!!!

## 5.1 Boxplots by `r pca_factor2` {.tabset .tabset-fade}

- All reported compounds plotted based on a factor

```{r, results='asis', echo = FALSE, include=T, warning=FALSE, fig.width=8, fig.height=6}


#Plot compounds by injection order
# Filter data
filter_wf = 
  filter_smeta(mode = "include",
               levels = inc_levels, 
               factor_name = inc_factor)


sample_data =  predicted(model_apply(filter_wf, importData))

sample_data = sample_data$data %>%
  mutate(treatment = sprintf("%s_%s", sample_data$sample_meta[[factor_bps1]], sample_data$sample_meta[[factor_bps2]])) %>%
  select(treatment, everything())

#Generate plots

for (compound in compound_names)   {
  
  compound_col = which(colnames(sample_data) == compound)
  
  axis.graphtitle.text <- element_text(face = "bold", color = "black", size = 12) #Label style
  axis.graph.text <- element_text(face = "bold", color = "black", size = 8)
  
  axis.limit.value <- quantile(sample_data[,compound_col], prob = 1, na.rm=TRUE)*1.3 # 40% extra limit for each compound. Quantiles are used to have most of the compounds on the graph. Extreme outliers may be out.
  
  
  #Create graph Sample ID
  not_detected = c()
  
  if(all(is.na(unlist(sample_data[,compound_col])))){
    not_detected  = c(not_detected, sprintf("%s not detected", compound))
  } else{
    
    p <- ggplot(sample_data, aes(`treatment`, unlist(sample_data[,compound_col]))) +
      theme_classic() +
      geom_boxplot(aes(fill = treatment), alpha = 0.1) +
      geom_point(aes(color = treatment), size = 4, alpha = 0.5) +
      ylim(c(0,axis.limit.value)) +
      labs(x="Treatment", y = units) + 
      theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1), legend.position="none") +
      #theme(legend.position="bottom", axis.text.x=element_text(angle = 0, hjust = 0.5, vjust = 0), axis.title = axis.graphtitle.text, axis.text = axis.graph.text) +
      ggtitle(compound)
    
    cat( paste('### ', compound, '\n' ) )
    print(p)
    cat('\n \n')
  }
}

print(not_detected)

```


## 5.2 Compound intensity plots {.tabset .tabset-fade}

```{r, results="asis", echo = FALSE, include=T, warning=FALSE, fig.width=10, fig.height=5}

#Plot compounds by injection order

# Filter data
filter_wf = 
    filter_smeta(mode = "include",
               levels = comp_int_levels, 
               factor_name = comp_int_factor)


injection_data =  predicted(model_apply(filter_wf, importData))

injection_data = injection_data$data %>%
  mutate(type = injection_data$sample_meta[[comp_int_id_header]],
         injection_order = injection_data$sample_meta$Injection_order) %>%
  select(type, injection_order, everything())


# Iterate over each compound

for (compound in compound_names)   {

  compound_col = which(colnames(injection_data) == compound)
  
  # Set axis limits
  axis.graphtitle.text <- element_text(face = "bold", color = "black", size = 12) #Label style
  axis.graph.text <- element_text(face = "bold", color = "black", size = 8)
  
  axis.limit.value <- quantile(injection_data[,compound_col], prob = 0.95, na.rm=TRUE)*1.2
  
  axis.limit.valueq75 <- quantile(injection_data[,compound_col], prob = 0.95, na.rm=TRUE)*1.2

  
  #Generate plot
    
  p <- ggplot(injection_data, aes(`injection_order`, unlist(injection_data[,compound_col]))) +
    theme_classic() +   
    geom_point(aes(color = type), size = 3, alpha = 0.6) +
    labs(x="Injection order", y = units) + 
    theme(legend.position="bottom", axis.text.x=element_text(angle = 0, hjust = 0.5, vjust = 0), 
          axis.title = axis.graphtitle.text, axis.text = axis.graph.text) +
      ggtitle(compound)
  
  
  cat( paste('### ', compound, '\n' ) )
  print(p)
  cat('\n \n')

}


```


# 6. Selection of samples

Finally, samples are selected as a combination of confirmed detection, presence in blanks and presence in at least 5 samples.

Different Tables are Exported to be included in the final report.


```{r}


# Select compounds to be included. Includes final filter by %CV and technical judgement

#compoundInfo_final<- compoundInfo %>% filter(Report == "YES") %>%
#                              select(-"Integration",-"our.coments",-"Included", -"Detected",
#                                      -"Include",-"Report",-"Use_type")


#compoundInfo_excluded <- compoundInfo %>% filter(Included == "YES", Report == "NO") %>% 
#                                       select(-"Integration",-"our.coments",-"Included", -"Detected",
#                                      -"Include",-"Report",-"Use_type")

#compound_excluded <- compoundInfo_excluded %>% select(Compound_name) %>% as.vector()

#compound_included<- compoundInfo_final %>% select(Compound_name) %>% as.vector()

#smetadata <- colnames(samples_file_replaced [1:35])

#finalfile_names_plasma <- samples_file_replaced %>% select(smetadata, compound_included[,1])



# Export files to merge in a the final report.
# This can be done with xlsx package for direct organization in Sheets but there are troubles with Java and Biomedicum computers
# Doing it with write_xlsx will be more practical (the files are edited later anyway)

#write_xlsx(finalfile_names_plasma, "FinalResults/ResultsFile_Replaced.xlsx") # The final datatable

#write_xlsx(compoundInfo_final, "FinalResults/ReportedCompoundMetadata_Replaced.xlsx") 
#write_xlsx(compoundInfo_excluded, "FinalResults/ReportedCompoundMetadataEx_Replaced.xlsx") 

#Save file for future R data processing

#list_selected_data <- list(finalfile_names_plasma = finalfile_names_plasma,
#                           compoundInfo_final = compoundInfo_final,
#                           compoundInfo_excluded = compoundInfo_excluded)

#saveRDS(list_selected_data, file = sprintf("%s/FinalData_Set_SL_Replaced.RDS", result_dir))


```


# 7. Final considerations {.tabset}

List of compounds reported and excluded from the final results

## 7.1. Compounds included

```{r}

#DT::datatable(compoundInfo_final,
#  caption = "Summary of Compounds included")

```

## 7.2 Compounds excluded

```{r}

#DT::datatable(compoundInfo_excluded,
#  caption = "Summary of Compounds included")

```



# 8. Session Information

Packages used for the creation of this file

```{r session, echo = FALSE, warning=FALSE}

sessionInfo()

```


